---
title: "assignment"
author: "Donny Liu"
date: "13/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Necessary Packages
```{r, echo=FALSE}
library(dplyr)
library(ggplot2)
```

Download file

The codes here will download the file, then unzip it, and reads the file into R
```{r}
fileUrl1 <- "https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2FNEI_data.zip"

if(!file.exists("./data")){
        dir.create("./data")
}

download.file(fileUrl1, destfile = "./data/pm.zip", method = "curl")
unzip("./data/pm.zip", exdir = ".")

NEI <- readRDS("summarySCC_PM25.rds")
SCC <- readRDS("Source_Classification_Code.rds")
```

We can quickly take a look at the two datasets
```{r}
View(NEI)
View(SCC)
```

Question 1

* We start by grouping the NEI dataset by year, summarize the emissions into anual then remove the NAs
```{r}
total_emissions_year <- NEI %>% 
        group_by(year) %>%
        summarize(Total.Emissions = sum(Emissions, na.rm = TRUE))

total_emissions_year
```

* Now we can plot it with the base plotting system
```{r}
with(total_emissions_year, 
     plot(x = year,
          y = Total.Emissions,
     ylab = "Total Annual Emissions in Tons",
     xlab = "Year",
     main = "Total Annual Emissions in the US by year",
     cex = 2,
     pch = 2,
     col = "red",
     lwd = 3))
```

* Then to find delta between 2008 and 1999
```{r}
total_emissions_2008 <- total_emissions_year[total_emissions_year$year == 2008, 2]
total_emissions_1999 <- total_emissions_year[total_emissions_year$year == 1999, 2]

delta_total_emissions <- total_emissions_2008 - total_emissions_1999
```

```{r}
View(delta_total_emissions)
```
        
        Total emissions from PM2.5 have decreased in the US from 1999 to 2008 by 3,868,761 tons also can be seen from the plot. 
        
Question 2

```{r}
total_emissions_baltimore <- NEI %>%
        subset(fips == "24510") %>%
        group_by(year) %>%
        summarize(Total.Emissions.Baltimore = sum(Emissions, na.rm = TRUE))

with(total_emissions_baltimore, 
     plot(x = year,
          y = Total.Emissions.Baltimore,
          ylab = "Total Annual Emissions in tons",
          xlab = "Year",
          main = "Total Annual Emissions in Baltimore by year",
          cex = 2,
          pch = 2,
          col = "red",
          lwd = 3))
```

```{r}
total_emissions_baltimore_2008 <- total_emissions_baltimore[total_emissions_baltimore$year == 2008, 2]

total_emissions_baltimore_1999 <- total_emissions_baltimore[total_emissions_baltimore$year == 1999, 2]

delta_total_emissions_baltimore <- total_emissions_baltimore_2008 - total_emissions_baltimore_1999
```

```{r}
View(delta_total_emissions_baltimore)
```

        Total emissions from PM2.5 in Baltimore have decreased from 1999 to 2008 by 1411.898 tons, as shown in the plot. 
        
Question 3

```{r}
total_emissions_type <- NEI %>%
        subset(fips == "24510") %>%
        group_by(year, type) %>%
        summarize(Total.Emissions.Type = sum(Emissions, na.rm = TRUE))

emissions_type <- ggplot(data = total_emissions_type, aes(year, Total.Emissions.Type))

emissions_type <- emissions_type +
        geom_point(color = "red",
                   size = 4,
                   alpha = 1/3) +
        facet_grid(. ~ type) +
        xlab("Year") +
        ylab("Total Emissions in tons") +
        ggtitle("Total Annual Emissions in Baltimore by Year")

emissions_type
```

        All source types appear to have decreased from 1999 to 2008 in Baltimore, except for the "point" type source. 
        
Question 4

* To solve this one, we should first merge the NEI and SCC datasets to see EL.Sector column which will show contribution from coal that is not due to combustion

```{r}
SCC_coal_comb <- SCC[grep("[Cc]oal", SCC$EI.Sector), ]
NEI_sub_coal <- subset(NEI,
                       NEI$SCC %in% SCC_coal_comb$SCC)

NEI_coal_comb <- merge(x = NEI_sub_coal,
                       y = SCC, 
                       by.x = "SCC",
                       by.y = "SCC")

NEI_coal_comb_tot <- NEI_coal_comb %>%
        group_by(year) %>%
        summarize(Total.Coal.Comb = sum(Emissions, na.rm = TRUE))

NEI_coal_comb_plot <- ggplot(NEI_coal_comb_tot, aes(year, Total.Coal.Comb))

NEI_coal_comb_plot <- NEI_coal_comb_plot +
        geom_point(color = "red",
                   size = 4,
                   alpha = 1/3) +
        xlab("Year") +
        ylab("Total Emissions in Tons") +
        ggtitle("Total Annual Coal Combustion Emissions in the US")

NEI_coal_comb_plot
```

```{r}
NEI_coal_comb_tot_2008 <- NEI_coal_comb_tot[NEI_coal_comb_tot$year == 2008, 2]
NEI_coal_comb_tot_1999 <- NEI_coal_comb_tot[NEI_coal_comb_tot$year == 1999, 2]

NEI_coal_comb_delta <- NEI_coal_comb_tot_2008 - NEI_coal_comb_tot_1999
```

```{r}
View(NEI_coal_comb_delta)
```

        Total coal combustion emissions in the US have decreased by 228,694.3 tons from 1999 to 2008. 
        
Question 5

```{r}
vehicle_scc <- SCC[grep("[Vv]eh", SCC$Short.Name), ]

emissions_motor_baltimore <- NEI %>%
        subset(fips == "24510" & NEI$SCC %in% vehicle_scc$SCC) %>%
        merge(y = vehicle_scc, by.x = "SCC", by.y = "SCC") %>%
        group_by(year) %>%
        summarize(Vehicle.Emissions.Type = sum(Emissions, na.rm = TRUE))

emissions_motor_baltimore_plot <- ggplot(emissions_motor_baltimore, aes(year, Vehicle.Emissions.Type)) + geom_point(color = "red", size = 4, alpha = 1/3) +
        xlab("Year") +
        ylab("Total Emissions in Tons") +
        ggtitle("Total Annual Vehicle Emissions in Baltimore City")

emissions_motor_baltimore_plot
```

```{r}
emissions_motor_baltimore_2008 <- emissions_motor_baltimore[emissions_motor_baltimore$year == 2008, 2]

emissions_motor_baltimore_1999 <- emissions_motor_baltimore[emissions_motor_baltimore$year == 1999, 2]

delta_motor_baltimore <- emissions_motor_baltimore_2008 - emissions_motor_baltimore_1999
```

```{r}
View(delta_motor_baltimore)
```

        Emissions from motor vehicles in Baltimore have decreased by 258.5445 tons from 1999 to 2008
        
Question 6

```{r}
vehicle_scc <- SCC[grep("[Vv]eh", SCC$Short.Name), ]

emissions_motor_la <- NEI %>%
        subset(fips == "06037" & NEI$SCC %in% vehicle_scc$SCC) %>%
        merge(y = vehicle_scc, by.x = "SCC", by.y = "SCC") %>%
        group_by(year) %>%
        summarize(Vehicle.Emissions.Type = sum(Emissions, na.rm = TRUE))

emissions_motor_baltimore2 <- cbind(emissions_motor_baltimore, "City" = rep("Baltimore", 4))
emissions_motor_la2 <- cbind(emissions_motor_la, "City" = rep("LA", 4))

emissions_motor_comp <- rbind(emissions_motor_baltimore2, emissions_motor_la2)

emissions_motor_comp_plot <- ggplot(emissions_motor_comp, aes(year, Vehicle.Emissions.Type, col = City)) +
        geom_point(size = 4, alpha = 1/3) +
        xlab("Year") +
        ylab("Total Emissions in Tons") +
        ggtitle("Comparison of Total Annual Vehicle Emissions between Baltimore and LA")

emissions_motor_comp_plot
```

```{r}
emissions_motor_la_2008 <- emissions_motor_la[emissions_motor_la$year == 2008, 2]
emissions_motor_la_1999 <- emissions_motor_la[emissions_motor_la$year == 1999, 2]

delta_la <- emissions_motor_la_2008 - emissions_motor_la_1999
View(delta_la)

abs(delta_la) > abs(delta_motor_baltimore)
```

        Los Angeles experienced an increase of emissions for 163. 44 tons, while the greatest magnitude change happened in Baltimore motor vehicle emissions, which was a decrease of 258.5445 tons between 1999 and 2008 









